\begin{thebibliography}{9}

\bibitem{cartea2015algorithmic}
Cartea, Álvaro, Sebastian Jaimungal, and José Penalva. \textit{Algorithmic and High-Frequency Trading}. Cambridge University Press, 2015.

\bibitem{abernethy2013adaptive}
Abernethy, Jacob, Yiling Chen, and Jennifer Wortman Vaughan. "Adaptive Market Making via Online Learning." In \textit{Proceedings of the 14th ACM Conference on Electronic Commerce}, 2013, pp. 365--382.

\bibitem{haarnoja2018soft}
Haarnoja, Tuomas, Aurick Zhou, Pieter Abbeel, and Sergey Levine. "Soft Actor-Critic: Off-Policy Maximum Entropy Deep Reinforcement Learning with a Stochastic Actor." In \textit{International Conference on Machine Learning (ICML)}, 2018, pp. 1861--1870.

\bibitem{espeholt2018impala}
Espeholt, Lasse, Hubert Soyer, Remi Munos, et al. "IMPALA: Scalable Distributed Deep-RL with Importance Weighted Actor-Learner Architectures." In \textit{Proceedings of the 35th International Conference on Machine Learning (ICML)}, 2018, pp. 1407--1416.

\bibitem{cont2010stochastic}
Cont, Rama, Arseniy Kukanov, and Sasha Stoikov. "The Price Impact of Order Book Events." \textit{Journal of Financial Econometrics} 12, no. 1 (2014): 47--88.

\bibitem{liang2018rllib}
Liang, Eric, Richard Liaw, Robert Nishihara, et al. "RLlib: Abstractions for Distributed Reinforcement Learning." In \textit{Proceedings of the 35th International Conference on Machine Learning (ICML)}, 2018, pp. 3053--3062.

\bibitem{zhang2021multi}
Zhang, Shangtong, and Shimon Whiteson. "Multi-Agent Reinforcement Learning: A Selective Overview of Theories and Algorithms." \textit{arXiv preprint arXiv:2106.11324}, 2021.

\end{thebibliography}